---
title: "R Notebook"
output: html_notebook
---

## 3.8 Position adjustments
- color a bar chart:
    - ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
- Each colored rectangle represents a combination of cut and clarity.
    - ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
- position = "fill" works like stacking, but makes each set of stacked bars the same height. 
- position = "dodge" places overlapping objects directly beside one another.

###3.8.1 Exercises
1).using a jitter position adjustment to decrease overplotting.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(position = "jitter")
```

2). width controls the amount of vertical displacement, and
height controls the amount of horizontal displacement.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point(position = position_jitter())
```

3).The geom geom_jitter() adds random variation to the locations points of the graph.The geom geom_count() sizes the points relative to the number of observations.
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = class)) +
  geom_jitter()
```


4).

```{r}
ggplot(data = mpg, aes(x = drv, y = hwy, colour = class)) +
  geom_boxplot()
ggplot(data = mpg, aes(x = drv, y = hwy, colour = class)) +
  geom_boxplot(position = "identity")
```

##3.9 Coordinate systems
- coord_flip() switches the x and y axes
    - ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
    geom_boxplot()
    - ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
    coord_flip()
- coord_quickmap() sets the aspect ratio correctly for maps. 
- coord_polar() uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.

###3.9.1 Exercises

1).
```{r}
ggplot(mpg, aes(x = factor(1), fill = drv)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y")
```
2).The labs function adds axis titles, plot titles, and a caption to the plot.
```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot() +
  coord_flip() +
  labs(
    y = "Highway MPG",
    x = "Class",
    title = "Highway MPG by car class",
    subtitle = "1999-2008",
    caption = "Source: http://fueleconomy.gov"
  )
```
3).The coord_quickmap() project is faster than coord_map() both because the projection is computationally easier, and unlike coord_map(), the coordinates of the individual geoms do not need to be transformed.

4).The function coord_fixed() ensures that the line produced by geom_abline() is at a 45-degree angle. 
A 45-degree line makes it easy to compare the highway and city mileage to the case in which city and highway MPG were equal.
```{r}
p <- ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() +
  geom_abline()
p + coord_fixed()
```

##3.10 The layered grammar of graphics

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
  
  - The seven parameters in the template compose the grammar of graphics, a formal system for building plots.
  - choose a geometric object to represent each observation in the transformed data
  - select a coordinate system to place the geoms into. 
  -You’d use the location of the objects to display the values of the x and y variables


#5 Data Transformation

###5.1.1 prerequisites
```{r}
library(nycflights13)
library(tidyverse)
```

###5.1.2 nycflights13

These describe the type of each variable:
- int stands for integers.
- dbl stands for doubles, or real numbers.
- chr stands for character vectors, or strings.
- dttm stands for date-times (a date + a time).

###5.1.3 dplyr basics

- Pick observations by their values (filter()).
- Reorder the rows (arrange()).
- Pick variables by their names (select()).
- Create new variables with functions of existing variables (mutate()).
- Collapse many values down to a single summary (summarise()).

All verbs work similarly:

- The first argument is a data frame.
- The subsequent arguments describe what to do with the data frame, using the variable names (without quotes).
- The result is a new data frame.

##5.2 Filter rows with filter()

- filter() allows you to subset observations based on their values. 


###5.2.1 comparisons

- R provides the standard suite: >, >=, <, <=, != (not equal), and == (equal).

filter(flights, month = 1)
#> Error: `month` (`month = 1`) must not be named, do you need `==`?

- There’s another common problem you might encounter when using ==: floating point numbers.

sqrt(2) ^ 2 == 2
#> [1] FALSE
1 / 49 * 49 == 1
#> [1] FALSE

- Instead of relying on ==, use near():
near(sqrt(2) ^ 2,  2)
#> [1] TRUE
near(1 / 49 * 49, 1)
#> [1] TRUE

###5.2.2 Logical operators

filter(flights, month == 11 | month == 12)
nov_dec <- filter(flights, month %in% c(11, 12))
filter(flights, !(arr_delay > 120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
- As well as & and |, R also has && and ||. Don’t use them here! You’ll learn when you should use them in conditional execution.

###5.2.3 missing values

- NA represents an unknown value so missing values are “contagious”: almost any operation involving an unknown value will also be unknown.

NA > 5
#> [1] NA
10 == NA
#> [1] NA
NA + 10
#> [1] NA
NA / 2
#> [1] NA

- to determine if a value is missing, use is.na():
is.na(x)
#> [1] TRUE

- filter() only includes rows where the condition is TRUE; it excludes both FALSE and NA values. 
- If you want to preserve missing values, ask for them explicitly:

###5.2.4 Exercise

1).
  a).
```{r}
filter(flights, arr_delay >= 120)
```
  
  b).
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
filter(flights, dest %in% c("IAH", "HOU"))
```
  
  c).
```{r}
airlines
filter(flights, carrier %in% c("AA", "DL", "UA"))
```
  
  d).
```{r}
filter(flights, month >= 7, month <= 9)
filter(flights, month %in% 7:9)
filter(flights, month == 7 | month == 8 | month == 9)
```
  
  e).
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```
  f).
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```
  
  g).
```{r}
summary(flights$dep_time)
filter(flights, dep_time <= 600 | dep_time == 2400)
c(600, 1200, 2400) %% 2400
filter(flights, dep_time %% 2400 <= 600)
```
2).The expression between(x, left, right) is equivalent to x >= left & x <= right.

```{r}
filter(flights, between(month, 7, 9))
```
3). arrival time (arr_time) is missing for these rows
```{r}
filter(flights, is.na(dep_time))
```

4). NA ^ 0 == 1 since for all numeric values x^0 = 1
NA | TRUE is TRUE because the value of the missing TRUE or FALSE, x or TRUE is TRUE for all values of x

###5.3.1 exercises

1). The arrange() function puts NA values last. To put NA values first, we can add an indicator of whether the column has a missing value

```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

2).The most delayed flight was HA 51, JFK to HNL, which was scheduled to leave on January 09, 2013 09:00. But the departure was delayed 1,301 minutes, which is 21 hours, 41 minutes. 

3).
```{r}
fastest_flights <- mutate(flights, mph = distance / air_time * 60)
fastest_flights <- select(
  fastest_flights, mph, distance, air_time,
  flight, origin, dest, year, month, day
)
```

4). The longest flight is HA 51, JFK to HNL, which is 4,983 miles.

```{r}
arrange(flights, desc(distance))
```

#5.4 Select columns with select()

- select() is not terribly useful with the flights data because we only have 19 variables
- select() in conjunction with the everything() helper.

###5.4.1
1).
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, contains("_time"), contains("arr_"))
```

2).The select() call ignores the duplication

3). The one_of() function selects variables with a character vector rather than unquoted variable name arguments.

- This function is useful because it is easier to programmatically generate character vectors with variable names than to generate unquoted variable names, which are easier to type.

4).The default behavior for contains() is to ignore case.
```{r}
select(flights, contains("TIME"))
select(flights, contains("TIME", ignore.case = FALSE))


